COG-ED Manuscript

### 2021-10-13 06_Paper/COG-ED/Manuscript.Rmd
update analysis

## Analysis
We conducted all analysis as described in Westbrook et al. [-@Westbrook2013].
The performance measure *d'* was computed as the difference of the *z*-transformed hit rate and the *z*-transformed false alarm rate [@Macmillan1990].
Declining performance was investigated by calculating an analysis of variance (ANOVA) with three linear contrasts comparing *d'* between two levels of 2-, 3-, and 4-back at a time.
Another ANOVA with three linear contrasts was computed to compare the mean reaction time (RT) between two levels of 2-, 3-, and 4-back at a time.
To investigate changes in NTLX ratings, six ANOVAs were computed, one for each NTLX subscale, and each with six linear contrasts comparing the ratings between two levels of 1-, 2-, 3-, and 4-back at a time.
The SVs were calculated by adding or subtracting the final iteration value of 0.015625 from the monetary value of the flexible level for each ED round, depending on the participant's choice in the last iteration.
Then, the final monetary values were divided by 2€, and the SV of each level per participant was computed by averaging all final monetary values of each level, regardless of whether it was fixed or flexible.
An ANOVA with six linear contrasts was computed, comparing the SVs between two levels of 1-, 2-, 3-, and 4-back at a time.
Each ANOVA in the analysis was followed up by a Tukey's test.
To determine the influence of task performance in the association of SVs and n-back level, we set up a multi-level model.
**MLM explanation goes here**
The association of ED and NFC was done using three separate regressions.
One regression used the AUC of each participant's SVs to predict their NFC score.
The second regression additionally included the mean of the NTLX subscales' AUCs of each participant as a predictor.
And the last regression used the AUC of each participant's SV, the AUC of the NTLX mental load subscale, and the AUC of the NTLX effort subscale to predict each participant's NFC score.

@@

  # ANOVA with three linear contrasts, contrasting the d’ of two n-back levels (2,3,4) at a time

@@

      h1a_data <- dprime[dprime$level != 1,]
      h1a_data$level <- as.factor(h1a_data$level)

@@

      hypothesis1a_anova <- summary(aov(h1a_data$d ~ h1a_data$level))
      hypothesis1a_posthoc <- TukeyHSD(aov(h1a_data$d ~ h1a_data$level))
      # TODO  check for posthoc test alternative
      #       anova calculation
      remove(h1a_data)

@@

# FIXME what about RT preprocessing
  # correct / postcorrect     *see Westbrook et al. 2013, table 1: correct trials
  # log-transformation
  # valid answers: false alarm vs. missing -> plausible responses

@@

@@

  data_hyp3 <- data.frame(subject = character(), auc_sv = double())

@@

    newdata <- data.frame(subject = data_SV$subject[subjectindex[i]],
                          auc_sv = auc(data_SV$level[subjectindex[i]:(subjectindex[i+1]-1)], data_SV$sv[subjectindex[i]:(subjectindex[i+1]-1)],
                                   method = "trapezoid", sort = FALSE))
    data_hyp3 <- rbind(data_hyp3, newdata)

@@

    nfc <- data_quest[setindex[i] + 2, grep("nfc", colnames(data_quest))]
    nfc[ ,c(4,6,7,8,9,10,11,12,15,16)] <-  nfc[ ,c(4,6,7,8,9,10,11,12,15,16)] * -1

@@

@@

  data_hyp3 <- merge(data_hyp3, data_nfc)

@@

  hypothesis3a <- summary(lm(data_hyp3$nfc ~ data_hyp3$auc_sv))

@@

  data_hyp3 <- merge(data_hyp3, ntlx)

@@

  hypothesis3b <- summary(lm(data_hyp3$nfc ~ data_hyp3$auc_sv + data_hyp3$auc_ntlx))

@@

  hypothesis3c <- summary(lm(data_hyp3$nfc ~ data_hyp3$auc_sv + data_hyp3$auc_mental + data_hyp3$auc_effort))

>>>>

## Analysis
We conducted all analysis as described in Westbrook et al. [-@Westbrook2013].
The performance measure *d'* was computed as the difference of the *z*-transformed hit rate and the *z*-transformed false alarm rate [@Macmillan1990].
Declining performance was investigated by calculating an analysis of variance (ANOVA) with three linear contrasts comparing *d'* between two levels of 2-, 3-, and 4-back at a time.
Another ANOVA with three linear contrasts was computed to compare the mean reaction time (RT) between two levels of 2-, 3-, and 4-back at a time.
To investigate changes in NTLX ratings, six ANOVAs were computed, one for each NTLX subscale, and each with six linear contrasts comparing the ratings between two levels of 1-, 2-, 3-, and 4-back at a time.
The SVs were calculated by adding or subtracting the final iteration value of 0.015625 from the monetary value of the flexible level for each ED round, depending on the participant's choice in the last iteration.
Then, the final monetary values were divided by 2€, and the SV of each level per participant was computed by averaging all final monetary values of each level, regardless of whether it was fixed or flexible.
An ANOVA with six linear contrasts was computed, comparing the SVs between two levels of 1-, 2-, 3-, and 4-back at a time.
Each ANOVA in the analysis was followed up by a Tukey's test.
# would it be possible to compare the originally calculated SV with our new one?
To determine the influence of task performance in the association of SVs and n-back level, we set up a multi-level model.
**MLM explanation goes here**
The association of ED and NFC was done using three separate regressions.
One regression used the AUC of each participant's SVs to predict their NFC score.
The second regression additionally included the mean of the NTLX subscales' AUCs of each participant as a predictor.
And the last regression used the AUC of each participant's SV, the AUC of the NTLX mental load subscale, and the AUC of the NTLX effort subscale to predict each participant's NFC score.

@@

  # ANOVA with three linear contrasts, contrasting the d’ of two n-back levels (2,3,4) at a time
  # FIXME:  actually, I guess, we have a rmANOVA here as we look at repeated measures of a subject's reaction time 
  #         under different load levels (= time points), see below

@@

      h1a_data <- dprime[dprime$level != 1,]      # FIXME: Westbrook et al. exclude level 1, do you understand why?
      h1a_data$level <- as.factor(h1a_data$level)

@@

      hypothesis1a_anova <- summary(aov(h1a_data$d ~ h1a_data$level))         # this calculates ANOVA without the specified contrasts and might be redundant
      hypothesis1a_anova.lm <- summary.lm(aov(h1a_data$d ~ h1a_data$level))   # this gives you the contrasted result
      hypothesis1a_posthoc <- TukeyHSD(aov(h1a_data$d ~ h1a_data$level))      # post hoc analyses should be redundant as we already contrasted the levels

  # FIXME:  rmANOVA
      library(ez) # compute repeated measurement ANOVA
      h1a_data$subject <- factor(h1a_data$subject)
      # if there are reasonable concerns that we have NAs within our data set, we need to exclude this data beforehand

      # https://stackoverflow.com/questions/35891062/planned-contrasts-using-ezanova-output-in-r
      # set orthogonal contrasts
      options(contrasts = c("contr.sum", "contr.poly"))
      hypothesis1a_rmanova <- ezANOVA(h1a_data, dv = d, wid = subject, within = level, type = 3, return_aov = T) 
      # I propose to use type-III sums of squares even if we do have a balanced design, but we may discuss that. 
      # alternatively we could use afex, results might be plotted easier; it is also possible to set up a linear model
      # TODO: find a way to plot means
        # if Mauchly's Test for Sphericity is sign > report Greenhouse-Geisser Epsilon (GGe) or Huynh-Feldt-Epsilon (HFe)

      # get emmeans
      library(emmeans)
      # this only works without the manually defined contrasts in line 439
      hypothesis1a_emm <- emmeans(hypothesis1a_rmanova$aov, ~ level) 
      # these means should be used for reporting/ plotting the means per load level

      # get the contrasts
      pairs(hypothesis1a_emm, adjust = "Holm")
      # or
      contrast(hypothesis1a_emm, list(c1 = c(1,-1,0), c2 = c(1,0,-1), c3 = c(0,1,-1)))

      # alternatively, post hoc tests can be run
      # as we have 3 levels in factor load level, we would run a fisher's least significant differences (FLSD)
      hypothesis1a_fisher <- ezStats(h1a_data, dv = d, wid = subject, within = level, type = 3)
      # calculate and save mean differences
      p1.p2 <- hypothesis1a_fisher$Mean[1] - hypothesis1a_fisher$Mean[2]
      p1.p3 <- hypothesis1a_fisher$Mean[1] - hypothesis1a_fisher$Mean[3]
      p2.p3 <- hypothesis1a_fisher$Mean[2] - hypothesis1a_fisher$Mean[3]

      #if absolute value is higher than fsld, the pairwise comparison is significant
      abs(p1.p2) > hypothesis1a_fisher$FLSD[1]
      abs(p1.p3) > hypothesis1a_fisher$FLSD[1]   
      abs(p2.p3) > hypothesis1a_fisher$FLSD[1]

      remove(h1a_data)

@@

# FIXME what about RT preprocessing?
  # correct / postcorrect     *see Westbrook et al. 2013, table 1: correct trials
  # log-transformation
  # valid answers: false alarm vs. missing -> plausible responses

@@

      # TODO: see H1a. 
      num_vars <- c("mental", "physical", "time", "performance", "effort", "frustration")
      h1c_data[num_vars] <- lapply(h1c_data[num_vars], as.numeric)
      hypothesis1c_rmanova <- ezANOVA(h1c_data, dv = mental, wid = subject, within = level, 
                                       type = 3, return_aov = T)
      #...

      # We may also run a rmMANOVA
      # https://cran.r-project.org/web/packages/MANOVA.RM/vignettes/Introduction_to_MANOVA.RM.html
      library(MANOVA.RM)
      # we need to convert our data frame to clean wide or long format

@@

  h3_data <- data.frame(subject = character(), auc_sv = double())

@@

    newdata <- data.frame(subject = data_SV$subject[subjectindex[i]],
                          auc_sv = auc(data_SV$level[subjectindex[i]:(subjectindex[i+1]-1)], data_SV$sv[subjectindex[i]:(subjectindex[i+1]-1)],
                                   method = "trapezoid", sort = FALSE))
    h3_data <- rbind(h3_data, newdata)

@@

    nfc <- data_quest[setindex[i] + 2, grep("nfc", colnames(data_quest))]
    # define and invert items to be recoded
    nfc[ ,c(4,6,7,8,9,10,11,12,15,16)] <-  nfc[ ,c(4,6,7,8,9,10,11,12,15,16)] * -1

@@

# further preprocessing of nfc-data ---------------------------------------------
  # FIXME: actually we would now
  # 1. determine deviation from univariate normality using Shapiro Wilk test
  # 2. normalize and standardize sum scores by Blom's formula (1954)
  # 3. again determine deviation from univariate normality of data from second step using Shapiro Wilk test
  # this would mean:

  # required scripts (might be best at the beginning, following the packages) --

  ## for normalizing variables
normalize.gen <- function(data, whichFormula=2, whichMeanSD=1) {
  # normalizes DATA using one of the formulae:
  #
  # 1 = Van der Waerden (1952): r * / (n + 1)
  # 2 = Blom (1954):           (r - 3/8) / (n + 1/4) (default)
  # 3 = Bliss (1956; Rankit):  (r - 1/2) / n
  # 4 = Tukey (1962):          (r - 1/3) / (n + 1/3)
  #
  # Ties will be treated as average, and the output will be rescaled
  # to have mean = 0 and sd = 1, if you enter 1 for mean0.sd1 (default);
  # if you want to keep the mean and sd of the original variable, enter 0.

  normalize = function(data,normalize.formula=whichFormula,mean0.sd1=whichMeanSD) {
    r = rank(data, na.last="keep", ties.method = ("average"))
    n = sum(!is.na(r))

    if (normalize.formula == 1)
      x = r/(n+1)
    if (normalize.formula == 2)
      x = (r-3/8)/(n+1/4)
    if (normalize.formula == 3)
      x = (r-1/2)/n
    if (normalize.formula == 4)
      x = (r-1/3)/(n+1/3)

    if (mean0.sd1 == 1)
      normal = qnorm(x, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)
    else
      normal = qnorm(x, mean = mean(data,na.rm = T), sd = sd(data,na.rm = T), lower.tail = TRUE, log.p = FALSE)
  }

  if (class(data)=="numeric") { 
    normal=normalize(data,whichFormula,whichMeanSD)
  } else {
    normal=NULL
    for (i in 1:dim(data)[2]) {
      normal=cbind(normal,normalize(data[,i],whichFormula,whichMeanSD))
    }
    colnames(normal)=names(data)
  }

  return(normal)

  ## wrapper for shapiro.test() 
sw <- function(x) {
  x = data.frame(x)
  d = dim(x)
  n = names(x)
  r = NULL
  for (i in 1:dim(x)[2]) {
    y = shapiro.test(x[,i])
    v = cbind(y$statistic,y$p.value)
    colnames(v) = c("W","p")
    rownames(v) = n[i]
    r = rbind(r,v)
  }
  return(round(r,3))
}

#------------------------------------------------------------------------------------

#determine deviation from univariate normality using Shapiro Wilk test
shapiro.summary = sw(data_nfc$nfc)

#normalize and standardize sum scores by Blom's formula (1954)
data_nfc$nnfc = data.frame(normalize.gen(data_nfc$nfc))
colnames(data_nfc$nnfc) = "nfc.normal"

#determine deviation from univariate normality using Shapiro Wilk test
shapiro.summary = rbind(shapiro.summary, sw(data_nfc$nnfc))

# end --------------------------------------------------------------------------

@@

  h3_data <- merge(h3_data, data_nfc)

@@

  hypothesis3a.lm <- summary(lm(h3_data$nfc ~ h3_data$auc_sv))

  # FIXME: additionally, we could calculate the multiple correlation between both variables
  hypothesis3a.mcorr <- sqrt(hypothesis3a.lm$r.squared)
  # this correlation seems surprisingly high ...

@@

  h3_data <- merge(h3_data, ntlx)

@@

  hypothesis3b <- summary(lm(h3_data$nfc ~ h3_data$auc_sv + h3_data$auc_ntlx))

@@

  hypothesis3c <- summary(lm(h3_data$nfc ~ h3_data$auc_sv + h3_data$auc_mental + h3_data$auc_effort))
